---
name: Annabelle
surname: Petherbridge
position: "Senior Financial Analyst"
phone: +1 (971)-277-0758
www: yourwebsite.com
email: "Annabelleelizabethh.com"
github: Annbelle123
linkedin: www.linkedin.com/in/annabelle-p-736040190
date: "`r format(Sys.time(), '%B %Y')`"
output: 
  vitae::latexcv:
    theme: classic
header-includes:
   - \usepackage{enumitem}
   - \setlist[itemize]{nosep, left=0pt}
   - \setlist[enumerate]{nosep, left=0pt}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

if (!requireNamespace(c("tinytex","vitae","tibble"), quietly = TRUE)) {
  install.packages(c("tinytex","vitae","tibble"))
}

if (Sys.which("pdflatex") == "" && !tinytex::is_tinytex()) {                    
  message("► TinyTeX not found – installing")
  tinytex::install_tinytex()        
} else {
  message("✓ LaTeX present – no installation needed.")
}

library(vitae)
library(tibble)
```

# Professional Summary

Analytical and results-driven finance professional with a Bachelor of Science in Finance and a minor in Economics from the University of Oregon. Currently serving as a Financial Analyst at HD Supply, with a strong track record in P&L management, forecasting, and cost analysis. Demonstrated ability to support cross-functional teams through data-driven insights and strategic recommendations. Previous experience includes roles in supply chain analysis and entrepreneurial ventures, showcasing a unique blend of corporate finance expertise and entrepreneurial acumen. Adept at leveraging Excel and financial tools to drive operational efficiency and cost savings. Passionate about advancing into financial planning and analysis leadership, with long-term goals in real estate investment and business development.

# Technical Skills

- **Programming Languages:** R, Python, SQL, etc.  
- **Web Technologies:** HTML, CSS/SCSS, Quarto, CMS, Google Analytics, etc.

# Education

```{r}
tribble(
  ~ Degree, ~ Year, ~ College, ~ Institution,
  "Master of [Your Field]", "20XX – Expected Graduation 20XX", "[Your School or Department]", "[Your University]",
  "Bachelor of [Your Field]", "20XX – 20XX", "[Your School or Department]", "[Your University]"
) %>%
  detailed_entries(Degree, Year, College, Institution)
```

# Work Experience

```{r}
tribble(
  ~ Years, ~ Company, ~ Duties, ~ Title,

  "Month YYYY – Month YYYY", "Company or Organization Name",
  list(
    "Summarize responsibility or achievement \\#1.",
    "Highlight tools, technologies, or methods used.",
    "Mention metrics, goals, or impact if possible.",
    "Add communication or leadership responsibilities here."
  ),
  "Job Title",

  "Month YYYY – Month YYYY", "Another Company or Team",
  list(
    "Led a project or team focused on [insert focus].",
    "Collaborated cross-functionally with [departments/roles].",
    "Managed content, campaigns, or dashboards.",
    "Contributed to strategic planning or reporting."
  ),
  "Job Title"

) %>%
  detailed_entries(
    what = Title,
    when = Years,
    with = Company,
    why = Duties
  )
```
