---
title: "Order Level Forecasting"
description: "A dashboard summarizing historical order trends and predicting future order volume"
categories: [Sales, Visualizations, Financial Analytics, Volume Analytics, Forecasting]
index: 2
---

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}

# Load Libraries
library(janitor) 
library(tidyverse) 
library(lubridate)
library(forecast)
library(shiny)
library(plotly)
library(dplyr)

# Load Data
data <- read.csv("../data/SuperStoreSales.csv")

# clean column names using janitor
data <- clean_names(data)

# check for missing values
data %>% 
  is.na() %>% 
  sum()

# pull records with missing values
sapply(data, function(x)which(is.na(x)))
data[c(2235,5275,8799,9147,9148,9149,9387,9388,9389,9390,9742),]

# na values driven by burlington city missing postal code. add postal code for burlington
data$postal_code<-replace_na(data$postal_code,05401)

# Check for remaining missing values
data %>% 
  is.na() %>% 
  sum()

# Convert dates to Date format
data$order_date <- as.Date(data$order_date, format = "%d/%m/%Y")
data$ship_date <- as.Date(data$ship_date, format = "%d/%m/%Y")

# Calculate summary stats
total_sales <- sum(data$sales, na.rm = TRUE)
num_orders <- n_distinct(data$order_id)
avg_sales_per_order <- round(mean(data$sales, na.rm = TRUE), 2)

# Format total sales using scales package
formatted_total_sales <- scales::dollar(total_sales)

# Comma-separated number of orders
formatted_num_orders <- scales::comma(num_orders)

 # Dollar sign for avg sales
formatted_avg_sales_per_order <- scales::dollar(avg_sales_per_order) 
```

<!-- CSS for card layout -->

```{=html}
<style>
  .card-container {
    display: flex;
    justify-content: space-around;
    gap: 20px;
    flex-wrap: wrap;
    width: 100%; /* Ensure the container takes up full width */
  }

  .card {
    border: 1px solid #ccc;
    padding: 15px;
    border-radius: 8px;
    width: 45%; /* Default width for cards */
    box-shadow: 0 4px 6px rgba(0, 0, 0, 0.1);
    margin-bottom: 20px;
  }

  /* Larger cards for specific sections like Sales Over Time and Top 3 Sales Months */
  .card.sales-time {
    width: 100% !important;  /* Ensure this card takes full width */
    min-height: 500px;  /* Increase height to accommodate the graph */
    padding: 20px;
    margin-bottom: 40px; /* Extra space for large cards */
  }

  .card.top-sales {
    width: 100% !important; /* Keep this card full width if needed */
    min-height: 400px;
    padding: 20px;
  }
  
  .card.product-segment {
  width: 100% !important;
  min-height: 400px;
  padding: 20px;
  margin-bottom: 40px;
}

  @media (min-width: 768px) {
    .card {
      width: 30%; /* 3 cards in a row for larger screens */
    }

    /* Keep the larger cards at 100% width for wide plots/tables */
    .card.sales-time {
      width: 100% !important;  /* Full width for Sales Over Time on larger screens */
    }

    .card.top-sales {
      width: 48%; /* Slightly larger for top-sales */
    }
  }

  @media (max-width: 600px) {
    .card {
      width: 100%; /* Stack vertically on mobile */
    }

    .card.sales-time {
      width: 100% !important;  /* Ensure full width on mobile */
      min-height: 300px;  /* Adjust for smaller screens */
    }
  }
</style>
```

# 🛒 About the Dataset

This dashboard is built using a sales dataset from Kaggle that captures **four years of transactions** from a global superstore. Each row represents a unique order, with details spanning **customer segments**, **order and shipping dates**, **shipping methods**, **product categories**, and **sales revenue**.

The dataset enables analysis of key business insights including:

-   **Sales trends over time**

-   **Product and category profitability**

-   **Customer behavior by segment**

# 📦 Order Volume Trends

-   Monthly order count over time.

-   Highlight seasonality and growth.


::: {.card .sales-time}
<h5 class="card-title">
📦 Monthly Order Volume
</h5>

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.height=6, fig.width=10}
library(ggplot2)

orders_by_month <- data %>%
  mutate(month = floor_date(order_date, "month")) %>%
  group_by(month) %>% 
  summarise(order_count = n_distinct(order_id))

ggplot(orders_by_month, aes(x = month, y = order_count)) +
  geom_line(color = "#2C3E50", linewidth = 1.2) +
  geom_point(color = "#2980B9", size = 2) +
  labs(title = "Monthly Order Volume",
       x = "Month",
       y = "Number of Orders") +
  theme_minimal()
```
:::


# 🔮 Forecasting Future Order Volume

::: {.card .sales-time}
<h5 class="card-title">
🔮 Forecasting Future Order Volume
</h5>

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.height=6, fig.width=10}
# Forecasting with Seasonality using Seasonal ARIMA (SARIMA)

# Prepare time series data
ts_data <- orders_by_month %>%
  arrange(month) %>%
  pull(order_count) %>%
  ts(start = c(year(min(orders_by_month$month)), month(min(orders_by_month$month))), frequency = 12)

# Fit Seasonal ARIMA model
fit <- auto.arima(ts_data, seasonal = TRUE)

# Forecast next 12 months
forecasted <- forecast(fit, h = 12)

# Plot forecast
autoplot(forecasted) +
  labs(title = "Forecasted Monthly Order Volume with Seasonality",
       x = "Month",
       y = "Predicted Orders") +
  theme_minimal()
```
:::

This plot represents the forecasted overall monthly order volume for the next 12 months using a Seasonal ARIMA (SARIMA) model. The historical data reveals a clear seasonal pattern with recurring peaks and troughs. The forecast projects this pattern forward, anticipating higher volumes during peak seasons and lower volumes during off-peak periods. The shaded confidence intervals indicate the expected range of future values, supporting strategic planning for demand surges and slowdowns.

# 📈  Forecasting Order Volume by Product Category

::: {.card .sales-time}
<h5 class="card-title">
📈  Forecasting Order Volume by Product Category
</h5>

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.height=6, fig.width=10}
# Forecasting Order Volume by Product Category

library(forecast)
library(ggplot2)
library(dplyr)
library(lubridate)
library(tidyr)

# Prepare monthly order count by category
category_monthly <- data %>%
  mutate(month = floor_date(order_date, "month")) %>%
  group_by(category, month) %>%
  summarise(order_count = n_distinct(order_id), .groups = "drop")

# Create a list of forecasts by category
category_forecasts <- category_monthly %>%
  group_by(category) %>%
  group_split() %>%
  setNames(unique(category_monthly$category)) %>%
  lapply(function(df) {
    ts_data <- ts(df$order_count,
                  start = c(year(min(df$month)), month(min(df$month))),
                  frequency = 12)
    fit <- auto.arima(ts_data, seasonal = TRUE)
    forecast(fit, h = 12)
  })

# Plot forecasts
par(mfrow = c(1, length(category_forecasts)))
for (cat in names(category_forecasts)) {
  plot(category_forecasts[[cat]],
       main = paste("Forecast for", cat),
       ylab = "Orders", xlab = "Month")
}
```
:::

This set of plots shows the forecasted order volume for each product category, also using SARIMA models. Each category reveals distinct seasonal behaviors:

-  🪑 Furniture: Steady growth with seasonal peaks around holidays.
-  🗂️ Office Supplies: Cyclical demand with spikes during back-to-school periods.
-  💻 Technology: Rapid growth with occasional surges tied to product launches and promotions.

These forecasts help tailor inventory and marketing strategies to the unique demand cycles of each category.

#  🧑‍💼 Forecasting Order Volume by Customer Segment

::: {.card .sales-time}
<h5 class="card-title">
  🧑‍💼 Forecasting Order Volume by Customer Segment
</h5>

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.height=6, fig.width=10}
# Forecasting Monthly Order Count by Segment with Seasonality

library(forecast)
library(dplyr)
library(lubridate)
library(tidyr)

# Prepare monthly order count by segment
segment_monthly <- data %>%
  mutate(month = floor_date(order_date, "month")) %>%
  group_by(segment, month) %>%
  summarise(order_count = n_distinct(order_id), .groups = "drop")

# Forecast and plot for each segment
segments <- unique(segment_monthly$segment)
par(mfrow = c(length(segments), 1))  # One row per segment

for (seg in segments) {
  df <- segment_monthly %>%
    filter(segment == seg) %>%
    complete(month = seq(min(month), max(month), by = "month"), fill = list(order_count = 0)) %>%
    arrange(month)

  ts_data <- ts(df$order_count,
                start = c(year(min(df$month)), month(min(df$month))),
                frequency = 12)

  fit <- auto.arima(ts_data, seasonal = TRUE)
  forecasted <- forecast(fit, h = 12)

  plot(forecasted,
       main = paste("Forecasted Monthly Orders -", seg, "Segment"),
       ylab = "Order Count", xlab = "Month")
}
```
:::

This set of plots presents a 12-month forecast of monthly order volume for each customer segment—Consumer, Corporate, and Home Office—using Seasonal ARIMA (SARIMA) models. Each segment exhibits unique seasonal patterns:

-  Consumer: Displays consistent seasonal peaks, often aligning with retail cycles and promotional periods.
-  Corporate: Shows steadier growth with moderate seasonal variation, reflecting business procurement cycles.
-  Home Office: Exhibits more irregular patterns, with occasional spikes likely tied to remote work trends or small business activity.

The shaded confidence intervals around each forecast line represent the expected range of future values, helping to guide segment-specific planning for inventory, staffing, and marketing efforts.

## 📎 Appendix

This dashboard was created using data from a public Kaggle dataset: [Sales Dataset on Kaggle](https://www.kaggle.com/datasets/rohitsahoo/sales-forecasting)
