---
title: "Actionable Insights on Forecasted Sales"
description: "A dashboard Analyzing Forecast Results and Actionable Insights on Predictions"
categories: [Sales, Visualizations, Financial Analytics, Volume Analytics, Forecasting]
index: 2
---

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}

# Load Libraries
library(janitor) 
library(tidyverse) 
library(lubridate)
library(forecast)
library(shiny)
library(plotly)
library(dplyr)
library(knitr)
library(scales)
library(tidyr)


# Load Data
data <- read.csv("../data/SuperStoreSales.csv")

# clean column names using janitor
data <- clean_names(data)

# check for missing values
data %>% 
  is.na() %>% 
  sum()

# pull records with missing values
sapply(data, function(x)which(is.na(x)))
data[c(2235,5275,8799,9147,9148,9149,9387,9388,9389,9390,9742),]

# na values driven by burlington city missing postal code. add postal code for burlington
data$postal_code<-replace_na(data$postal_code,05401)

# Check for remaining missing values
data %>% 
  is.na() %>% 
  sum()

# Convert dates to Date format
data$order_date <- as.Date(data$order_date, format = "%d/%m/%Y")
data$ship_date <- as.Date(data$ship_date, format = "%d/%m/%Y")

# Calculate summary stats
total_sales <- sum(data$sales, na.rm = TRUE)
num_orders <- n_distinct(data$order_id)
avg_sales_per_order <- round(mean(data$sales, na.rm = TRUE), 2)

# Format total sales using scales package
formatted_total_sales <- scales::dollar(total_sales)

# Comma-separated number of orders
formatted_num_orders <- scales::comma(num_orders)

 # Dollar sign for avg sales
formatted_avg_sales_per_order <- scales::dollar(avg_sales_per_order) 
```

<!-- CSS for card layout -->

```{=html}
<style>
  .card-container {
    display: flex;
    justify-content: space-around;
    gap: 20px;
    flex-wrap: wrap;
    width: 100%; /* Ensure the container takes up full width */
  }

  .card {
    border: 1px solid #ccc;
    padding: 15px;
    border-radius: 8px;
    width: 45%; /* Default width for cards */
    box-shadow: 0 4px 6px rgba(0, 0, 0, 0.1);
    margin-bottom: 20px;
  }

  /* Larger cards for specific sections like Sales Over Time and Top 3 Sales Months */
  .card.sales-time {
    width: 100% !important;  /* Ensure this card takes full width */
    min-height: 500px;  /* Increase height to accommodate the graph */
    padding: 20px;
    margin-bottom: 40px; /* Extra space for large cards */
  }

  .card.top-sales {
    width: 100% !important; /* Keep this card full width if needed */
    min-height: 400px;
    padding: 20px;
  }
  
  .card.product-segment {
  width: 100% !important;
  min-height: 400px;
  padding: 20px;
  margin-bottom: 40px;
}

  @media (min-width: 768px) {
    .card {
      width: 30%; /* 3 cards in a row for larger screens */
    }

    /* Keep the larger cards at 100% width for wide plots/tables */
    .card.sales-time {
      width: 100% !important;  /* Full width for Sales Over Time on larger screens */
    }

    .card.top-sales {
      width: 48%; /* Slightly larger for top-sales */
    }
  }

  @media (max-width: 600px) {
    .card {
      width: 100%; /* Stack vertically on mobile */
    }

    .card.sales-time {
      width: 100% !important;  /* Ensure full width on mobile */
      min-height: 300px;  /* Adjust for smaller screens */
    }
  }
</style>
```

# 🛒 About the Dataset

This dashboard is built using a sales dataset from Kaggle that captures **four years of transactions** from a global superstore. Each row represents a unique order, with details spanning **customer segments**, **order and shipping dates**, **shipping methods**, **product categories**, and **sales revenue**.

The dataset enables analysis of key business insights including:

-   **Sales trends over time**

-   **Product and category profitability**

-   **Customer behavior by segment**


# ✅ 12 Month Sales Forecast Trended

::: {.card .sales-time}
<h5 class="card-title">

</h5>

```{r transposed-forecast-table, echo=FALSE, message=FALSE, warning=FALSE}

# Prepare monthly sales data
sales_by_month <- data %>%
  mutate(month = floor_date(order_date, "month")) %>%
  group_by(month) %>%
  summarise(sales_volume = sum(sales, na.rm = TRUE), .groups = "drop")

# Convert to time series
ts_data <- ts(sales_by_month$sales_volume,
              start = c(year(min(sales_by_month$month)), month(min(sales_by_month$month))),
              frequency = 12)

# Fit ARIMA model
fit <- auto.arima(ts_data, seasonal = TRUE)

# Forecast next 12 months
forecasted <- forecast(fit, h = 12)

# Create forecast table
forecast_table <- data.frame(
  Month = format(seq(max(sales_by_month$month) %m+% months(1), by = "month", length.out = 12), "%b.%Y"),
  `High Estimate` = forecasted$upper[,1],
  `Expected Forecast` = forecasted$mean,
  `Low Estimate` = forecasted$lower[,1]
)

# Transpose and format as currency
transposed_table <- t(forecast_table[,-1])
colnames(transposed_table) <- forecast_table$Month
rownames(transposed_table) <- c("High Estimate", "Expected Forecast", "Low Estimate")
transposed_table <- apply(transposed_table, c(1,2), function(x) dollar(x, accuracy = 1))

# Display the transposed table
kable(transposed_table, caption = "📅 12-Month Sales Forecast")
```
:::

This transposed table presents a 12-month forecast of total monthly sales using an ARIMA model. It includes high, expected, and low estimates for each month, offering a clear and concise view of projected revenue trends. The format supports quick comparisons across months and helps stakeholders plan for best- and worst-case scenarios.

# 🗂️ 12-Month Sales Forecast by Product Category

::: {.card .sales-time}
<h5 class="card-title">

📊 Forecast Tables for Furniture, Office Supplies, and Technology

</h5>

```{r 12monthfcstbyprodcat, echo=FALSE, message=FALSE, warning=FALSE}

# Prepare monthly sales data by category
category_monthly_sales <- data %>%
  mutate(month = floor_date(order_date, "month")) %>%
  group_by(category, month) %>%
  summarise(sales_volume = sum(sales, na.rm = TRUE), .groups = "drop")

# Function to forecast and format table for a category
forecast_category_table <- function(df, category_name) {
  df <- df %>%
    filter(category == category_name) %>%
    complete(month = seq(min(month), max(month), by = "month"), fill = list(sales_volume = 0)) %>%
    arrange(month)

  ts_data <- ts(df$sales_volume,
                start = c(year(min(df$month)), month(min(df$month))),
                frequency = 12)

  fit <- auto.arima(ts_data, seasonal = TRUE)
  forecasted <- forecast(fit, h = 12)

  forecast_table <- data.frame(
    Month = format(seq(max(df$month) %m+% months(1), by = "month", length.out = 12), "%b.%Y"),
    `High Estimate` = forecasted$upper[,1],
    `Expected Forecast` = forecasted$mean,
    `Low Estimate` = forecasted$lower[,1]
  )

  transposed_table <- t(forecast_table[,-1])
  colnames(transposed_table) <- forecast_table$Month
  rownames(transposed_table) <- c("High Estimate", "Expected Forecast", "Low Estimate")
  transposed_table <- apply(transposed_table, c(1,2), function(x) dollar(x, accuracy = 1))

  kable(transposed_table, caption = paste("📅 12-Month Sales Forecast -", category_name))
}

# Generate tables for each category
forecast_category_table(category_monthly_sales, "Furniture")
forecast_category_table(category_monthly_sales, "Office Supplies")
forecast_category_table(category_monthly_sales, "Technology")

```
:::

These three transposed tables provide detailed 12-month sales forecasts for each product category—Furniture, Office Supplies, and Technology. Each table includes high, expected, and low estimates, allowing for category-specific planning and performance tracking. This breakdown supports more granular decision-making in inventory, marketing, and budgeting.

# 📈 Actionable Insights

::: {.card .sales-time}
<h5 class="card-title">

📈 Actionable Insights

</h5>

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Function to get top 10 products for a given category
get_top_products <- function(category_name) {
  data %>%
    filter(category == category_name) %>%
    group_by(product_name) %>%
    summarise(total_sales = sum(sales, na.rm = TRUE), .groups = "drop") %>%
    arrange(desc(total_sales)) %>%
    slice_head(n = 10) %>%
    mutate(total_sales = dollar(round(total_sales, 0))) %>%
    kable(caption = paste("🏆 Top 10 Products by Sales –", category_name))
}

# Generate tables for each category
get_top_products("Furniture")
get_top_products("Office Supplies")
get_top_products("Technology")
```
:::

# 📊 Financial Forecast Summary

::: {.card .sales-time}
### 🪑 Furniture – 12-Month Forecast Summary

- **Total High Estimate:** $294,317  
- **Total Expected Forecast:** $235,055  
- **Total Low Estimate:** $175,790  
- **📈 Peak Month:** November 2019  
- **📉 Lowest Month:** February 2019  

**Implications:**  
Furniture shows moderate growth potential with a strong Q4 peak. Inventory planning should ramp up ahead of November. February may be ideal for cost-saving initiatives or maintenance.

**Staffing Note:**  
Consider scaling fulfillment and delivery staff in Q4. Use February for training or cross-departmental support.
:::

::: {.card .sales-time}
### 🖇️ Office Supplies – 12-Month Forecast Summary

- **Total High Estimate:** $383,764  
- **Total Expected Forecast:** $281,772  
- **Total Low Estimate:** $179,783  
- **📈 Peak Month:** December 2019  
- **📉 Lowest Month:** February 2019  

**Implications:**  
Office Supplies show strong seasonal demand, especially in December. This category is highly responsive to B2B and institutional cycles.

**Staffing Note:**  
Prepare for a surge in packaging and customer service in Q4. February may allow for reduced shifts or reallocation to digital support.
:::

::: {.card .sales-time}
### 💻 Technology – 12-Month Forecast Summary

- **Total High Estimate:** $437,816  
- **Total Expected Forecast:** $299,571  
- **Total Low Estimate:** $161,328  
- **📈 Peak Month:** November 2019  
- **📉 Lowest Month:** February 2019  

**Implications:**  
Technology is the top-performing category across all forecast scenarios. High-value items drive strong Q4 performance, especially in November.

**Staffing Note:**  
Expand technical support and B2B sales teams ahead of Q4. February may be used for product training or customer retention campaigns.
:::

::: {.card .sales-time}
### 📈 Overall Forecast Insights

- **Best Performing Category:** Technology (Highest total forecast across all scenarios)
- **Worst Case Total Sales (All Categories, Low Estimate):** $516,901  
- **Expected Total Sales (All Categories):** $816,398  
- **Best Case Total Sales (All Categories, High Estimate):** $1,115,897  

**Strategic Takeaways:**
- Q4 (Oct–Dec) is critical across all categories—plan inventory, staffing, and marketing accordingly.
- February is consistently the weakest month—ideal for cost-saving, training, or system upgrades.
- Technology should be prioritized for investment, marketing, and enterprise sales focus.
:::


# 📊 Confidence-Based Actionable Insights

::: {.card .sales-time}
### 🪑 Furniture – Forecast-Based Actions

**Upper Bound (High Sales)**  
- Increase inventory of top sellers like *HON Task Chairs* and *Riverside Bookcases*.  
- Expand marketing and sales outreach to corporate clients.  
- Staffing: Hire or reassign additional warehouse staff and delivery personnel. Consider temporary staffing for peak months.

**Expected Forecast**  
- Maintain current stock levels and marketing cadence.  
- Focus on fulfillment efficiency and customer satisfaction.  
- Staffing: Keep staffing steady; cross-train employees to handle multiple roles.

**Lower Bound (Low Sales)**  
- Delay large inventory purchases and reduce marketing spend.  
- Offer targeted promotions to stimulate demand.  
- Staffing: Reduce overtime or shift hours. Reallocate staff to stronger categories or customer service.
:::

::: {.card .sales-time}
### 🖇️ Office Supplies – Forecast-Based Actions

**Upper Bound (High Sales)**  
- Stock up on binding machines and consumables.  
- Launch B2B subscription programs for recurring orders.  
- Staffing: Add fulfillment staff for packaging and shipping. Increase customer support for B2B clients.

**Expected Forecast**  
- Maintain steady supply and pricing.  
- Focus on customer retention and upselling.  
- Staffing: Maintain current staffing levels. Use part-time or flexible staff to manage fluctuations.

**Lower Bound (Low Sales)**  
- Reduce inventory of slow-moving SKUs.  
- Promote digital alternatives or bundle offers.  
- Staffing: Consolidate roles or reduce hours in fulfillment. Shift staff to marketing or digital sales support.
:::

::: {.card .sales-time}
### 💻 Technology – Forecast-Based Actions

**Upper Bound (High Sales)**  
- Prioritize inventory for high-value items like *Canon Copiers* and *Cisco TelePresence*.  
- Offer premium service packages and enterprise solutions.  
- Staffing: Hire or contract technical staff for installations and support. Expand B2B account teams.

**Expected Forecast**  
- Maintain balanced inventory and support resources.  
- Focus on upselling accessories and services.  
- Staffing: Maintain current staffing. Provide training to improve cross-selling and support.

**Lower Bound (Low Sales)**  
- Delay restocking expensive items.  
- Promote trade-in or upgrade programs.  
- Staffing: Freeze hiring and reduce hours for technical support. Reassign staff to retention or education roles.
:::

# 📎 Appendix

This dashboard was created using data from a public Kaggle dataset: [Sales Dataset on Kaggle](https://www.kaggle.com/datasets/rohitsahoo/sales-forecasting)

